A Data-Driven Framework for Neural Field Modelling

### 2010-11-19 ltx/first_order_framework/LaTeX_Manuscript.tex
miror corrections

\begin{keyword}
	keyword1 \sep keyword2
%% keywords here, in the form: keyword \sep keyword

@@

where $\left[ \boldsymbol\Psi(\mathbf{r}')\right]_{:i}$ denotes the $i^{th}$ column of $\boldsymbol{\Psi}(\mathbf{r}')$ which is a $n_x \times n_{\theta}$ matrix, where $n_x$ is the number of basis functions (and states) and $n_{\theta}$ is the number of connectivity kernel basis functions. This matrix is defined analytically (see \ref{ap:ConvOfGaussians} for analytic convolution of two Gaussians), since it is constant with respect to the states and the unknown parameters. Substituting equation~\ref{eq:DefPsi} into equation~\ref{eq:ReducedForm} gives

@@

\subsection{State and Parameter Estimation}\label{StateAndParameterEstimationSection} In this section, we describe the procedure for estimating the states, $\mathbf{x}_t$, the connectivity kernel parameters, $\boldsymbol \theta$, and the synaptic dynamics, $\xi$. The estimation process is a two part iterative algorithm, consisting of a state estimation step followed by a parameter estimation step. Once initialized, at each iteration the sequence of estimated state vectors is used to update the parameter set. The resulting parameters are then used when estimating the new state vector sequence for the next iteration. The procedure stops when the parameters converge. The algorithm is initialized using a bounded random state vector sequence, guaranteeing that the initial estimated parameter set forms a stable kernel.

The additive form of the unscented Rauch-Tung-Striebel smoother (URTSS)~\cite{Sarkka2010} was used for the state estimation. The URTSS incorporates an unscented Kalman filter (UKF)~\cite{Julier1997, Merwe2003} in a forward step to estimate posterior states, $\hat{\mathbf x}_t^{f}$, followed by a backward step to compute the smoothed state estimates, $\hat{\mathbf x}_t^{b}$. The first and second order moments of the predicted state are captured by propagating the so-called sigma points through the state equation. The sigma points, $\mathcal X_i$, are calculated using the unscented transform as follows:
\begin{align}\label{eq:sigmapoints1}
	\mathcal X_{0}&=\bar x \\
	\mathcal X_{i}&=\bar x+\left(\sqrt{( n_x + \lambda)\mathbf P_x}\right)_i, \quad i=1, \dots, n_x \\
	\mathcal X_{i}&=\bar x-\left(\sqrt{( n_x + \lambda)\mathbf P_x}\right)_{i- n_x}, \quad i= n_x+1, \dots, 2n_x 
\end{align}
where $\bar x$ represents either $\hat{\mathbf x}_t^{f}$ or $\hat{\mathbf x}_t^{b}$, $\mathbf{P}_x$ is the corresponding covariance matrix for filtering or smoothing, $\left(\sqrt{( n_x + \lambda)\mathbf P_x}\right)_i$ is the $i^{th}$ column of the scaled matrix square root of $\mathbf P_x$, and $n_x$ is the dimension of the state space. The total number of sigma points is $2n_x+1$. The scaling parameter, $\lambda$, is defined as 
\begin{equation}\label{eq:sigmapoints3}
	\lambda=\alpha^2( n_x+\kappa) - n_x, 
\end{equation}
where the constant $\alpha$ determines the spread of the sigma points and $\kappa$ is an additional scaling parameter (see~\cite{Haykin2001} for more details).

@@

Although the system is nonlinear, the parameters of the system are linear with respect to the state. This is exploited by our procedure where the parameter estimation uses a least squares (LS) method that minimizes the sum of the squared errors (of a predicted state update) with each new estimate of the state vector sequence. To create the least squares parameter estimator, we first define the $n_x \times n_{\theta}$ matrix
\begin{equation}
	\mathbf{q}(\mathbf{x}_t) = \int_\Omega \boldsymbol{\Psi}(\mathbf{r}') f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) \,\textrm{d}\mathbf{r}'.
\end{equation}
Now given a state estimate sequence from the initialisation or an iteration of the URTSS, we can write
\begin{align*}
	 \hat{\mathbf x}^b_{1} &= \mathbf{q}(\hat{\mathbf x}^b_0) \boldsymbol{\theta}+\xi\hat{\mathbf x}^b_0+\mathbf e_0 \\
	\hat{\mathbf x}^b_{2} &= \mathbf{q}(\hat{\mathbf x}^b_1) \boldsymbol{\theta}+\xi\hat{\mathbf x}^b_1+\mathbf e_1  \\
	&\vdots& \\
	\hat{\mathbf x}^b_{T}&= \mathbf{q}(\hat{\mathbf x}^b_{T-1}) \boldsymbol{\theta}+\xi\hat{\mathbf x}^b_{T-1}+\mathbf e_{T-1}. 
\end{align*}
This can be written in the compact form
\begin{equation}
	\mathbf Z=\mathbf X \mathcal W+\mathbf{e}, 
\end{equation}
where
\begin{small}
\begin{equation*}
	\mathbf Z=\left[
	\begin{array}{cccc}
		\hat{\mathbf x}^b_{1}\\
		\hat{\mathbf x}^b_{2}\\
		\vdots\\
		\hat{\mathbf x}^b_{T}
	\end{array}
	\right],\quad \mathbf X=\left[
	\begin{array}{cccc}
		\mathbf q(\hat{\mathbf x}^b_0)& \hat{\mathbf x}^b_{0}\\
		\mathbf q(\hat{\mathbf x}^b_1)& \hat{\mathbf x}^b_{1}\\
		\vdots & \vdots\\
		\mathbf q(\hat{\mathbf x}^b_{T-1})& \hat{\mathbf x}^b_{T-1}
	\end{array}
	\right] 
\end{equation*}
\end{small}
and
\begin{small}
\begin{equation*}
\quad \mathcal W=\left[
	\begin{array}{cc}
		\boldsymbol{\theta} \\
		\xi
	\end{array}
	\right],\quad \mathbf{e}=\left[
	\begin{array}{cccc}
		\mathbf e_0\\
		\mathbf e_1\\
		\vdots\\
		\mathbf e_{T-1}
	\end{array}
	\right].
\end{equation*}
\end{small}
Following this, the LS parameter estimates~\cite{Kailath2000}, $ \hat{\mathcal{W}}$, are
\begin{equation}
	\hat{\mathcal{W}}=(\mathbf X^\top\mathbf X)^{-1}\mathbf X^\top\mathbf Z. 
\end{equation}

>>>>

\begin{keyword}
	\parham{keyword1 \sep keyword2}
%% keywords here, in the form: keyword \sep keyword

@@

where $\left[ \boldsymbol\Psi(\mathbf{r}')\right]_{:i}$ denotes the $i^{th}$ column of $\boldsymbol{\Psi}(\mathbf{r}')$ which is a $n_x \times n_{\theta}$ matrix, where $n_x$ is the number of basis functions (and states) and $n_{\theta}$ is the number of connectivity kernel basis functions. This matrix is defined analytically (see \ref{ap:ConvOfGaussians} for analytic convolution of two Gaussians), since it is constant with respect to the states and the unknown parameters. \parham{Substituting $\boldsymbol{\Psi}(\mathbf{r}')$ into equation~\ref{eq:ReducedForm} gives}

@@

\subsection{State and Parameter Estimation}\label{StateAndParameterEstimationSection} In this section, we describe the procedure for estimating the states, $\mathbf{x}_t$, the connectivity kernel parameters, $\boldsymbol \theta$, and the synaptic dynamics, $\xi$. The estimation process is a two part iterative algorithm, consisting of a state estimation step followed by a parameter estimation step. Once initialized, at each iteration the sequence of estimated state vectors is used to update the parameter set. The resulting parameters are then used when estimating the new state vector sequence for the next iteration. The procedure stops when the \parham{parameter estimates} converge. The algorithm is initialized using a bounded random state vector sequence, guaranteeing that the initial estimated parameter set forms a stable kernel.

The additive form of the unscented Rauch-Tung-Striebel smoother (URTSS)~\cite{Sarkka2010} was used for the state estimation. The URTSS incorporates an unscented Kalman filter (UKF)~\cite{Julier1997, Merwe2003} in a forward step to estimate posterior states, $\hat{\mathbf x}_t^{f}$, followed by a backward step to compute the smoothed state estimates, $\hat{\mathbf x}_t^{b}$. The first and second order moments of the predicted state are captured by propagating the so-called sigma points through the state equation. The sigma points, $\mathcal X_i$, are calculated using the unscented transform as follows:
\begin{align}\label{eq:sigmapoints1}
	\mathcal X_{0}&=\mathbf{\bar x} \\
	\mathcal X_{i}&= \mathbf{\bar x}+\left(\sqrt{( n_x + \lambda)\mathbf P_x}\right)_i, \quad i=1, \dots, n_x \\
	\mathcal X_{i}&=\mathbf{\bar x}-\left(\sqrt{( n_x + \lambda)\mathbf P_x}\right)_{i- n_x}, \quad i= n_x+1, \dots, 2n_x 
\end{align}
where $\mathbf{\bar x}$ represents either $\hat{\mathbf x}_t^{f}$ or $\hat{\mathbf x}_t^{b}$, $\mathbf{P}_x$ is the corresponding covariance matrix for filtering or smoothing, $\left(\sqrt{( n_x + \lambda)\mathbf P_x}\right)_i$ is the $i^{th}$ column of the scaled matrix square root of $\mathbf P_x$, and $n_x$ is the dimension of the state space. The total number of sigma points is $2n_x+1$. The scaling parameter, $\lambda$, is defined as 
\begin{equation}\label{eq:sigmapoints3}
	\lambda=\alpha^2( n_x+\kappa) - n_x, 
\end{equation}
where the constant $\alpha$ determines the spread of the sigma points around $\mathbf{\bar x}$, and $\kappa$ is an additional scaling parameter (see~\cite{Haykin2001} for more details).

@@

Although the system is nonlinear, the parameters of the system are linear with respect to the state. This is exploited by our procedure where the parameter estimation uses a least squares (LS) method that minimizes the sum of the squared errors (of a predicted state update) with each new estimate of the state vector sequence. To create the least squares parameter estimator, we first define the $n_x \times n_{\theta}$ matrix
\begin{equation}
	\mathbf{q}(\mathbf{x}_t) = \int_\Omega \boldsymbol{\Psi}(\mathbf{r}') f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) \,\textrm{d}\mathbf{r}'.
\end{equation}
Now given a state estimate sequence from the initialisation or an iteration of the URTSS, we can write
\begin{align*}
	 \hat{\mathbf x}^b_{1} &= \mathbf{q}(\hat{\mathbf x}^b_0) \boldsymbol{\theta}+\xi\hat{\mathbf x}^b_0+\mathbf e_0 \\
	\hat{\mathbf x}^b_{2} &= \mathbf{q}(\hat{\mathbf x}^b_1) \boldsymbol{\theta}+\xi\hat{\mathbf x}^b_1+\mathbf e_1  \\
	&\vdots& \\
	\hat{\mathbf x}^b_{T}&= \mathbf{q}(\hat{\mathbf x}^b_{T-1}) \boldsymbol{\theta}+\xi\hat{\mathbf x}^b_{T-1}+\mathbf e_{T-1}. 
\end{align*}
This can be written in the compact form
\begin{equation}
	\mathbf Z=\mathbf X \mathcal W+\mathbf{e}, 
\end{equation}
where
\begin{small}
\begin{equation*}
	\mathbf Z=\left[
	\begin{array}{cccc}
		\hat{\mathbf x}^b_{1}\\
		\hat{\mathbf x}^b_{2}\\
		\vdots\\
		\hat{\mathbf x}^b_{T}
	\end{array}
	\right],\quad \mathbf X=\left[
	\begin{array}{cccc}
		\mathbf q(\hat{\mathbf x}^b_0)& \hat{\mathbf x}^b_{0}\\
		\mathbf q(\hat{\mathbf x}^b_1)& \hat{\mathbf x}^b_{1}\\
		\vdots & \vdots\\
		\mathbf q(\hat{\mathbf x}^b_{T-1})& \hat{\mathbf x}^b_{T-1}
	\end{array}
	\right] 
\end{equation*}
\end{small}
and
\begin{small}
\begin{equation*}
\quad \mathcal W=\left[
	\begin{array}{cc}
		\boldsymbol{\theta} \\
		\xi
	\end{array}
	\right],\quad \mathbf{e}=\left[
	\begin{array}{cccc}
		\mathbf e_0\\
		\mathbf e_1\\
		\vdots\\
		\mathbf e_{T-1}
	\end{array}
	\right].
\end{equation*}
\end{small}
Following this, the LS parameter estimates~\cite{Ljung1985}, $ \hat{\mathcal{W}}$, are
\begin{equation}
	\hat{\mathcal{W}}=(\mathbf X^\top\mathbf X)^{-1}\mathbf X^\top\mathbf Z. 
\end{equation}
